---
title: "longread_analysis"
author: "Jullien Flynn"
date: '2019-03-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This file goes through the analysis of long read data (PacBio and Nanopore) using the k-Seek + Phobos method.  


# 1) D. virilis PacBio data  

```{r}

Dvir_reps <-  read.delim("/Users/jullienflynn/Documents/AndyLab/Heterochromatin/PacBio_scripts/Revised_scripts/D_virilis_satellites_analysis_backup/LongRead_kseek_Phobos/ForR/D_virilis.all.phobos.simple.out", header=FALSE)

colnames (Dvir_reps) <- c("kmer", "count")

# summarize the sequences
Dvir_reps.simple <- tapply(Dvir_reps$count, Dvir_reps$kmer, sum)
total_bp <- 40935421853
genome_size <- 389000000
depth <- total_bp/genome_size

Dvir_reps_norm <- Dvir_reps.simple/depth

Dvir_reps_sorted_norm <- sort(Dvir_reps_norm, decreasing = T)

Dvir_reps_sorted_norm[1:5]
kmers_of_interest <- c("AAACTAC", "AAACTAT", "AAATTAC", "AAACAAC", "AAACGAC")
Dvir_reps_sorted_norm[kmers_of_interest]
sum(Dvir_reps_sorted_norm[kmers_of_interest])

```

# 2) D. novamexicana PacBio data  

```{r}
Dnova.simple.out.all <- read.delim("/Users/jullienflynn/Documents/AndyLab/Heterochromatin/PacBio_scripts/Revised_scripts/D_virilis_satellites_analysis_backup/LongRead_kseek_Phobos/ForR/Dnovamex.all.simple.out", header=FALSE)
colnames (Dnova.simple.out.all) <- c("kmer", "count")

# summarize the sequences
Dnova.simple <- tapply(Dnova.simple.out.all$count, Dnova.simple.out.all$kmer, sum)

Dnova_total_bases <- 40712554490
Dnova_genome_size <- 2.44E8
(Dnova_depth <- Dnova_total_bases/2.5E8)

Dnova_norm <- Dnova.simple/Dnova_depth

Dnova_sorted_norm <- sort(Dnova_norm, decreasing = T)

Dnova_sorted_norm[1:5]

Dnova_sorted_norm[kmers_of_interest]
sum(Dnova_sorted_norm[kmers_of_interest])


```

# 3) D. americana PacBio data  

```{r}

D_americana_reps <-  read.delim("/Users/jullienflynn/Documents/AndyLab/Heterochromatin/PacBio_scripts/Revised_scripts/D_virilis_satellites_analysis_backup/LongRead_kseek_Phobos/ForR/amCCS.all.phobos.simple.out", header=FALSE)

colnames (D_americana_reps) <- c("kmer", "count")

# summarize the sequences
D_americana_reps.simple <- tapply(D_americana_reps$count, D_americana_reps$kmer, sum)
D_americana_reps_sorted <- sort(D_americana_reps.simple, decreasing = T)

total_bp <- 2808871904
genome_size <- 240000000
(depth <- total_bp/genome_size)

D_americana_reps_sorted_norm <- D_americana_reps_sorted/depth

D_americana_reps_sorted_norm[1:5]
D_americana_reps_sorted_norm[kmers_of_interest]
sum(D_americana_reps_sorted_norm[kmers_of_interest])

```

# 4) D. hydei PacBio data 

```{r}

Dhyd.simple.out.all <- read.delim("/Users/jullienflynn/Documents/AndyLab/Heterochromatin/PacBio_scripts/Revised_scripts/D_virilis_satellites_analysis_backup/LongRead_kseek_Phobos/ForR/Dhydei.simple.out.all.filtered", header=FALSE)
colnames (Dhyd.simple.out.all) <- c("kmer", "count")

# summarize the sequences
Dhyd.simple <- tapply(Dhyd.simple.out.all$count, Dhyd.simple.out.all$kmer, sum)

assembly_size <- 1.77E8
reads_bases <- 19564840690
(depth <- reads_bases/assembly_size)

Dhyd.simple.norm <- Dhyd.simple/depth
Dhyd.simple.norm.sorted <- sort (Dhyd.simple.norm, decreasing = T)
Dhyd.simple.norm.sorted [1:10]


```

# 5) Now make a comparative plot  

```{r}
length_classify <- function (kmer_sequence) {
  x <- NULL
  if (nchar(kmer_sequence) == 1) { x <- "1mer" }
  else if (nchar(kmer_sequence) == 2) { x <- "2mer" }
  else if (nchar(kmer_sequence) > 2 & nchar(kmer_sequence) < 6) { x <- "3-5mer" }
  else if (nchar(kmer_sequence) == 6) { x <- "6mer" }
  else if (nchar(kmer_sequence) == 7) { x <- "7mer" }
  else if (nchar(kmer_sequence) > 7 & nchar(kmer_sequence) < 21) { x <- "8-20mer" }
  return (x)
}

sum_kmer_class <- function (abundance_vec, class_vector) {
  a <- sum(abundance_vec[which(class_vector == "1mer" )])
  b <- sum(abundance_vec[which(class_vector == "2mer" )])
  c <- sum(abundance_vec[which(class_vector == "3-5mer" )])
  d <- sum(abundance_vec[which(class_vector == "6mer" )])
  e <- sum(abundance_vec[which(class_vector == "7mer" )])
  f <- sum(abundance_vec[which(class_vector == "8-20mer" )])
  
  output <- c(a,b,c,d,e,f)
  names(output) <- c("1mer", "2mer", "3-5mer", "6mer", "7mer", "8-20mer")
  return(output)
}

Dvir_kmer_class <- sapply(names(Dvir_reps_sorted_norm), length_classify)
Dvir_sum_class <- sum_kmer_class(Dvir_reps_sorted_norm, Dvir_kmer_class)
Dvir_df <- data.frame(kmer_class=names(Dvir_sum_class), repeat_span=Dvir_sum_class, species="vir")

Dnova_kmer_class <- sapply(names(Dnova_sorted_norm), length_classify)
Dnova_sum_class <- sum_kmer_class(Dnova_sorted_norm, Dnova_kmer_class)
Dnova_df <- data.frame(kmer_class=names(Dnova_sum_class), repeat_span=Dnova_sum_class, species="nova")

Damer_kmer_class <- sapply(names(D_americana_reps_sorted_norm), length_classify)
Damer_sum_class <- sum_kmer_class(D_americana_reps_sorted_norm, Damer_kmer_class)
Damer_df <- data.frame(kmer_class=names(Damer_sum_class), repeat_span=Damer_sum_class, species="amer")

Dhyd_kmer_class <- sapply(names(Dhyd.simple.norm.sorted), length_classify)
Dhyd_sum_class <- sum_kmer_class(Dhyd.simple.norm.sorted, Dhyd_kmer_class)
Dhyd_df <- data.frame(kmer_class=names(Dhyd_sum_class), repeat_span=Dhyd_sum_class, species="hydei")

repeatspan_df <- rbind(Dvir_df, Dnova_df, Damer_df, Dhyd_df)
options(scipen=-1)
library(ggplot2)
ggplot(repeatspan_df, aes(fill=species, x=factor(kmer_class), y=repeat_span)) + 
  geom_bar(position="dodge", stat="identity") + 
  ylab("Total repeat span (bp)") +
  #scale_x_discrete(limits = repeatspan_df$kmer_class) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, size=13, hjust = 1)) +
  theme(axis.text.y = element_text(size=13), axis.title = element_text(size=13))

```

# 6) D. virilis Nanopore data  

```{r}

Dvir_pore_reps <-  read.delim("/Users/jullienflynn/Documents/AndyLab/Heterochromatin/PacBio_scripts/Revised_scripts/D_virilis_satellites_analysis_backup/LongRead_kseek_Phobos/ForR/Dvir.nanopore.phobos.all.simple.out", header=FALSE)

colnames (Dvir_pore_reps) <- c("kmer", "count")

Dvir_porereps.simple <- tapply(Dvir_pore_reps$count, Dvir_pore_reps$kmer, sum)

Dvir_porereps_sorted <- sort(Dvir_porereps.simple, decreasing = T)

reads_bases=7223297550
genome_size=3.89E8
(depth=reads_bases/genome_size)
Dvir_porereps_sorted_norm <- Dvir_porereps_sorted/depth

get_kmer_length <- function (kmer) {
  return(nchar(kmer))
}

Dvir_porereps_sorted_norm[which(get_kmer_length(names(Dvir_porereps_sorted_norm)) == 7)][1:10]

kmers_of_interest2 <- c("AAACTAC", "AAACTAT", "AAATTAC", "AAACAAC", "AAATCAC", "AGCCTAT")

Dvir_porereps_sorted_norm[kmers_of_interest2]
sum(Dvir_porereps_sorted_norm[kmers_of_interest2])

# make a plot now including Nanopore and PacBio amounts

vir_sat <- read.csv("/Users/jullienflynn/Documents/AndyLab/Heterochromatin/PacBio_scripts/Revised_scripts/D_virilis_satellites/LongRead_kseek_Phobos/ForR/virilis_satellite_quant.csv", header=T)
vir_sat

(Pacbio_sats <- subset(vir_sat, Platform=="PacBio"))
library(ggplot2)
ggplot(Pacbio_sats, aes(fill=Approach, y=Amount, x=Satellite)) +
  geom_bar(position="dodge", stat="identity") +
  theme_bw() +
  scale_x_discrete(limits = Pacbio_sats$Satellite) +
  scale_y_continuous(limits=c(0,2.6e+07)) +
  theme(legend.title=element_blank()) + 
  theme(text = element_text(size=14)) + 
  ylab("Amount (bp)") 

# this one is for my presentation
ggplot(Pacbio_sats, aes(fill=Approach, y=Amount, x=Satellite)) +
  geom_bar(position="dodge", stat="identity") +
  theme_bw() +
  scale_x_discrete(limits = Pacbio_sats$Satellite) +
  theme(legend.title=element_blank()) + 
  theme(text = element_text(size=14)) + 
  ylab("Amount (bp)")


(Pacbio_sats <- subset(vir_sat, Platform=="PacBio"))
library(ggplot2)
ggplot(Pacbio_sats, aes(fill=Approach, y=Amount, x=Satellite)) +
  geom_bar(position="dodge", stat="identity") +
  theme_bw() +
  scale_x_discrete(limits = Pacbio_sats$Satellite) +
  scale_y_continuous(limits=c(0,2.85e+07)) +
  theme(legend.title=element_blank()) + 
  theme(text = element_text(size=14)) + 
  ylab("Amount (bp)")

(Nanopore_sats <- subset(vir_sat, Platform=="Nanopore"))

ggplot(Nanopore_sats, aes(fill=Approach, y=Amount, x=Satellite)) +
  geom_bar(position="dodge", stat="identity") +
  theme_bw() +
  scale_x_discrete(limits = Nanopore_sats$Satellite) +
  scale_y_continuous(limits=c(0,2.85e+07)) +
  theme(legend.title=element_blank()) + 
  theme(text = element_text(size=14)) + 
  ylab("Amount (bp)") 

```

# 6) Strand bias in the PacBio and Nanopore data

```{r}
strand_bias <- read.csv(file="/Users/jullienflynn/Documents/AndyLab/Heterochromatin/PacBio_scripts/Revised_scripts/D_virilis_satellites/LongRead_kseek_Phobos/ForR/Strand_bias_summary2.csv", header=T)

Pacbio_strand <- subset(strand_bias, Platform=="PacBio_real")

ggplot(Pacbio_strand) +
  geom_bar(aes(y=Percent, x=Satellite, fill=factor(Strand, levels=c("neg", "pos"))), stat="identity") +
  theme_bw() +
  scale_x_discrete(limits = Pacbio_strand$Satellite) +
  theme(legend.title=element_blank()) +
  theme(text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1))

Pacbio_sim <- subset(strand_bias, Platform=="PacBio_sim")


ggplot(Pacbio_sim) +
  geom_bar(aes(y=Percent, x=Satellite, fill=factor(Strand, levels=c("neg", "pos"))), stat="identity") +
  theme_bw() +
  scale_x_discrete(limits = Pacbio_sim$Satellite) +
  theme(legend.title=element_blank()) +
  theme(text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1))


Nanopore_real <- subset(strand_bias, Platform=="Nanopore_real")

ggplot(Nanopore_real) +
  geom_bar(aes(y=Percent, x=Satellite,  fill=factor(Strand, levels=c("neg", "pos"))), stat="identity", width = 0.5) +
  theme_bw() +
  scale_x_discrete(limits = Nanopore_real$Satellite) +
  theme(legend.title=element_blank()) +
  theme(text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1))  

