---
title: "Polyteny_analysis"
author: "Jullien Flynn"
date: '2019-03-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script goes through the analysis to determine the effect of polyteny on quantification of satellites. First it uses D. melanogaster data: embryos and salivary glands from Yarosh and Spradling 2004, and flies from Gutzwiller et al. 2015.  

Then we go on to analyze D. virilis data we produced ourselves from imaginal discs from male larvae compared to male flies. We also include analysis before and after filtering here to show the bias against satellite containing reads.  

# 1) D. melanogaster analysis  

```{r}

kseek_data <- read.csv("/Users/jullienflynn/Documents/AndyLab/Heterochromatin/PacBio_scripts/Revised_scripts/D_virilis_satellites/Polyteny/ForR/melpolyteny.rep.compiled.GCcorr")

# clean up data
file_names <- as.vector(kseek_data$X)
rownames(kseek_data) <- file_names
# now get rid of the column that has the file name
kseek_data[,1] <- NULL
# the columns (kmers) by their mean abundance
means <- colMeans(kseek_data)
index.orders <- order(means, decreasing=T)
kseek_data_ordered <- kseek_data[,order(means, decreasing=T)]
# get rid of N 
kseek_data_ordered$N <- NULL

kmers_oi <- c("AACATAGAAT", "AATAT", "AATAG", "AAGAG", "AAGAC", "AACAC", "ACATATAT", "AAGAT", "AAGAGAG")

kmers_oi <- c("AACATAGAAT", "AAGAG", "AAGAC", "AACAC", "AAGAT", "AAGAGAG")

kseek_subset_koi <- data.frame(kseek_data_ordered[, kmers_oi])

kmer_abundances <- c()
kmer_names <- c()
sample_names <- c()
sample_category <- c()
library(stringr)
for (j in 1:ncol(kseek_subset_koi)) {
  for (i in 1:nrow(kseek_subset_koi)) {
    kmer_abundances <- c(kmer_abundances, kseek_subset_koi[i,j])
    kmer_names <- c(kmer_names, colnames(kseek_subset_koi)[j])
    x <- rownames(kseek_subset_koi)[i]
    y <- str_sub(x, 1, str_length(x)-1)
    sample_names <- c(sample_names, x)
    sample_category <- c(sample_category, y)
  }
}
 
kseek_to_plot <- data.frame (sample_names, sample_category, kmer_names, kmer_abundances )

library(ggplot2)

ggplot(data = kseek_to_plot, aes(x = kmer_names, y = kmer_abundances, fill = sample_category )) +
  geom_bar(stat = "summary", fun.y = "mean", position = "dodge") +
  geom_errorbar(stat = "summary", fun.data = mean_se, position = "dodge") +
  xlab("kmer") +
  ylab("Copy number") +
  theme (axis.text.x = element_text(angle=45, hjust=1))

# make a function that calculates the mean of all 3 groups for a specific kmer

calculate_group_means <- function(kmer) {
  embryo <- subset(kseek_to_plot, kmer_names==kmer & sample_category=="embryo")
  mean_embryo <- mean(embryo$kmer_abundances)
  fly <- subset(kseek_to_plot, kmer_names==kmer & sample_category=="fly")
  mean_fly <- mean(fly$kmer_abundances)
  salivary <- subset(kseek_to_plot, kmer_names==kmer & sample_category=="salivary")
  mean_salivary <- mean(salivary$kmer_abundances)
  result <- c(mean_embryo, mean_fly, mean_salivary)
  names(result) <- c("embryo", "fly", "salivary")
  return(result)
}

calculate_group_means("AAGAG")
calculate_group_means("AACATAGAAT")
calculate_group_means("AACAC")
calculate_group_means("AAGAC")

# look at microsatellites that are euchromatic now

kmers_oi <- c("AC", "AGC", "C")

kseek_subset_koi <- data.frame(kseek_data_ordered[, kmers_oi])

kmer_abundances <- c()
kmer_names <- c()
sample_names <- c()
sample_category <- c()

for (j in 1:ncol(kseek_subset_koi)) {
  for (i in 1:nrow(kseek_subset_koi)) {
    kmer_abundances <- c(kmer_abundances, kseek_subset_koi[i,j])
    kmer_names <- c(kmer_names, colnames(kseek_subset_koi)[j])
    x <- rownames(kseek_subset_koi)[i]
    y <- str_sub(x, 1, str_length(x)-1)
    sample_names <- c(sample_names, x)
    sample_category <- c(sample_category, y)
  }
}
 
kseek_to_plot <- data.frame (sample_names, sample_category, kmer_names, kmer_abundances )

ggplot(data = kseek_to_plot, aes(x = kmer_names, y = kmer_abundances, fill = sample_category )) +
  geom_bar(stat = "summary", fun.y = "mean", position = "dodge") +
  geom_errorbar(stat = "summary", fun.data = mean_se, position = "dodge") +
  xlab("kmer") +
  ylab("Copy number") +
  theme (axis.text.x = element_text(angle=45, hjust=1))  

```

# 2) D. virilis - before filtering  

```{r}
depth <- c(15.506, 3.894)
names(depth) <- c("Dvir_disks", "Dvir_flies")

data2 <- read.delim ('/Users/jullienflynn/Documents/AndyLab/Heterochromatin/PacBio_scripts/Revised_scripts/D_virilis_satellites/Polyteny/ForR/Dvir_polyteny.rep.compiled')
kmer.labels <- sapply(strsplit(colnames(data2), "\\."), '[', 1)
colnames(data2) <- kmer.labels
samples <- as.vector(data2$lines)
samples2 <- sapply(strsplit(samples, "[.]"), '[', 1)
data2[,1] <- NULL
rownames(data2) <- samples2

# now normalize the data
data2_normalized <- data2/depth

# now sort the data
kmer_means <- colMeans(data2)
data2_normalized_sorted <- data2_normalized[,order(kmer_means, decreasing=T)]

data2_normalized_sorted[,1:10]

# use the embryos - make a data frame so I can use it later.
kmers_of_interest <- c("AAACTAC", "AAACTAT", "AAATTAC")
post_qfiltering_flies <- data2_normalized_sorted[1, kmers_of_interest]

data2_normalized_sorted[, kmers_of_interest]
# total Mb of repeats
(sum(data2_normalized_sorted[1, kmers_of_interest])*7)/10^6
(sum(data2_normalized_sorted[2, kmers_of_interest])*7)/10^6

```

# 3) D. virilis - after filtering  

```{r}
depth <- c(18.44, 4.597)
names(depth) <- c("Dvir_disks", "Dvir_flies")
data3 <- read.delim ('/Users/jullienflynn/Documents/AndyLab/Heterochromatin/PacBio_scripts/Revised_scripts/D_virilis_satellites/Polyteny/ForR/Dvir_polyteny.allreads.rep.compiled')

kmer.labels <- sapply(strsplit(colnames(data3), "\\."), '[', 1)
colnames(data3) <- kmer.labels
samples <- as.vector(data3$lines)
samples2 <- sapply(strsplit(samples, "[.]"), '[', 1)
data3[,1] <- NULL
rownames(data3) <- samples2

data3[,1:10]
# now normalize the data
data3_normalized <- data3/depth

# now sort the data
kmer_means <- colMeans(data3)
data3_normalized_sorted <- data3_normalized[,order(kmer_means, decreasing=T)]

data3_normalized_sorted[,1:10]

no_qfiltering <- data3_normalized_sorted[1, kmers_of_interest]
no_qfiltering
rownames (no_qfiltering) <- c("no-filter")
rownames (post_qfiltering_flies) <- c("q-filter")
post_qfiltering_flies
filter_compare <- rbind(no_qfiltering, post_qfiltering_flies)


# now let's make a plot.
kmer_abundances <- c()
kmer_names <- c()
sample_category <- c()
for (j in 1:ncol(filter_compare)) {
  for (i in 1:nrow(filter_compare)) {
    kmer_abundances <- c(kmer_abundances, (filter_compare[i,j]*7)/10^6)
    kmer_names <- c(kmer_names, colnames(filter_compare)[j])
    sample_category <- c(sample_category, rownames(filter_compare)[i])
  }
}

filter_compare_Mb <- data.frame(sample_category, kmer_names, kmer_abundances)
filter_compare_Mb

library(ggplot2)
# Comparing fly data before and after filtering
ggplot(data = filter_compare_Mb, aes(x = kmer_names, y = kmer_abundances, fill = sample_category )) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw()

## combine with polyteny comparison

data_subset <- rbind(data3_normalized_sorted[c(1,2), kmers_of_interest], post_qfiltering_flies) 
data_subset

kmer_abundances <- c()
kmer_names <- c()
sample_names <- c()
sample_category <- c()
library(stringr)
# look at this in ggplot
for (j in 1:ncol(data_subset)) {
  for (i in 1:nrow(data_subset)) {
    kmer_abundances <- c(kmer_abundances, data_subset[i,j]*7)
    kmer_names <- c(kmer_names, colnames(data_subset)[j])
    x <- rownames(data_subset)[i]
    y <- substr(x, 6, str_length(x))
    sample_names <- c(sample_names, x)
    sample_category <- c(sample_category, y)
  }
}

polyteny_fig <- data.frame(sample_names, sample_category, kmer_names, kmer_abundances)
polyteny_fig

wei <- data.frame( "sample_names" = c("Dvir_wei", "Dvir_wei", "Dvir_wei"), "sample_category" = c("wei", "wei", "wei"), "kmer_names" = c("AAACTAC", "AAACTAT", "AAATTAC"), "kmer_abundances" = c(7099097, 490963, 540644))
str(wei)
wei

merged_polyteny <- rbind (polyteny_fig, wei)
merged_polyteny
merged_polyteny$kmer_abundances <- merged_polyteny$kmer_abundances/10^6

#install.packages("wesanderson")
library(wesanderson)

#options(scipen=999)
# darjeeling2 looks good to me.
ggplot(data = merged_polyteny, aes(x = kmer_names, y = kmer_abundances, fill = sample_category )) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  scale_fill_manual(values = wes_palette("Darjeeling2", n = 4)) +
  ylab ("Satellite amount (Mb)") +
  theme(text = element_text(size=12))

  
# total Mb of repeats
# discs
(sum(data3_normalized_sorted[1, kmers_of_interest])*7)/10^6
# flies
(sum(data3_normalized_sorted[2, kmers_of_interest])*7)/10^6

# looking at microsatellites, they are equivalent.
data3_normalized_sorted[ c(1,2), c("AT", "AC", "AG")]
# make a plot for the microsatellites

microsats <- c("AT", "AC", "AG")
data_subset <- data3_normalized_sorted[c(1,2), microsats]


kmer_abundances <- c()
kmer_names <- c()
sample_names <- c()
sample_category <- c()
library(stringr)
# look at this in ggplot
for (j in 1:ncol(data_subset)) {
  for (i in 1:nrow(data_subset)) {
    kmer_abundances <- c(kmer_abundances, data_subset[i,j]*2)
    kmer_names <- c(kmer_names, colnames(data_subset)[j])
    x <- rownames(data_subset)[i]
    y <- substr(x, 6, str_length(x))
    sample_names <- c(sample_names, x)
    sample_category <- c(sample_category, y)
  }
}

polyteny_fig <- data.frame(sample_names, sample_category, kmer_names, kmer_abundances)
polyteny_fig

ggplot(data = polyteny_fig, aes(x = kmer_names, y = kmer_abundances, fill = sample_category )) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() + theme(legend.title=element_blank()) + 
  theme(text = element_text(size=16)) + 
  ylab("Amount (bp)")


# I am  interested to see if the ratio is similar for these top 3 7mers. Answer: yes!
data3_normalized_sorted[2, kmers_of_interest]/data3_normalized_sorted[1, kmers_of_interest]
